The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-05-09 paper/ms.tex
changes from last telecon

\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{makecell}

@@

\newcommand{\sgra}{Sgr~A\textsuperscript{*}}

@@

\title{The Galactic Midplane Is Not a Plane: Implications for Dynamical
Analysis with \textit{Gaia} Data and Beyond}

% \correspondingauthor{Angus Beane}
\email{abeane@sas.upenn.edu}

\author[0000-0002-8658-1453]{Angus Beane}
\affil{\cca}
\affil{\penn}

\author[0000-0003-3939-3297]{Robyn Sanderson}
\affil{\cca}
\affil{\penn}

@@

\author[0000-0001-8275-9181]{Douglas Grion Filho}
\affil{\columbia}

\author[0000-0003-0064-4060]{Mordecai-Mark Mac Low}
\affil{\cca}
\affil{\amnh}

@@

\author[0000-0003-2866-9403]{David W. Hogg}
\affil{\nyuphys}
\affil{\nyucds}
\affil{\cca}
\affil{\mpia}

\author{Chervin F. P. Laporte}
\altaffiliation{CITA National Fellow}
\affil{\victoria}

@@

Stellar actions, computed for stars from their six-dimensional phase space
measurements and an assumed Galactic potential, are used to label and
distinguish orbits. In principle, actions have the attractive quality of being
invariant and thus good labels of a star's orbit. Such analyses are often used
to interpret data from the \textit{Gaia} mission. However, inaccurate
measurements of the phase space position and uncertainties in the assumed
Galactic potential induce an error in the computed action. We show that, in
addition to these complications, a systematic bias in the phase space position
induces a phase-dependence in the actions, which we interpret as a systematic
error. An offset in the vertical position of $\sim15\,\pc$ for a thin disk
orbit or $\sim120\,\pc$ for a thick disk orbit is sufficient to induce a
$25\,\%$ systematic error in the vertical action $J_z$ computed under the
assumption of an axisymmetric potential. The induced error distribution is
non-Gaussian and bimodal, with neither of the modes peaking on the null value.
Furthermore, we show that the local midplane varies by $\sim100\,\pc$ at the
Solar circle in Milky Way-mass cosmological zoom-in simulations from the FIRE
project. From observations that the mean vertical velocity of the Milky Way
disk varies by $\sim5\,\kms$ with radius, we estimate that the true midplane
variations are of order $\sim60\textup{--}90\,\pc$. Thus, current
state-of-the-art action calculations --- which assume the global and local
midplanes are the same --- are likely to include a systematic vertical offset
depending on the volume considered. Variation in the local standard of rest
induces similar issues. The variation of the midplane must be taken into
account when performing dynamical modeling across the large regions of the
disk accessible to \textit{Gaia} and future missions.

@@

\keywords{Galaxy: disk -- Galaxy: evolution -- Galaxy: kinematics and dynamics
-- Galaxy: structure -- solar neighborhood -- stars: kinematics and dynamics}

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution as a
result of \textit{Gaia} Data Release 2 (DR2).  Recent major discoveries
include the remnants of a major merger \citep{2018ApJ...860L..11K,
2018Natur.563...85H, 2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space
``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly
indicating local substructure infall \citep{2018MNRAS.481.1501B,
2018arXiv180800451L}, and a gap
suggestive of perturbation by a dark matter subtructure in the tidal stream
GD1 \citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
indicate that the Milky Way's stellar distribution, which demonstrably departs
significantly from axisymmetry, is undergoing phase mixing and dynamical
interactions across a range of spatial and temporal scales.

@@

Once a Galactocentric coordinate system has been established and a
six-dimensional (6D) phase space measurement of a star has been made, it is
often desirable to convert this measurement into action space to concisely
summarize its projected orbit, model the stellar distribution function, or
find stars with similar dynamical properties. Actions are conserved quantities
that describe the orbit of a star under the assumption of regular, bound
orbits in a system where the equations of motion are separable in a particular
coordinate system. They are the cyclical integral of the canonical momentum
over its conjugate position:
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{.}
\eeq
Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical,
and azimuthal coordinates respectively in a cylindrical coordinate system and
the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric
potential, these actions are invariant and $J_{\phi} \equiv L_z$, where $L_z$
is the $z$-component of the angular momentum
\citep{2008gady.book.....B,2014RvMP...86....1S}.

With the advent of 6D phase-space measurements over a relatively large
($\gtrsim 2$ kpc) volume from the \textit{Gaia} satellite, the study of stars'
actions has gained new popularity. One reason is dimensionality reduction---an
individual stellar orbit is concisely described by three actions, as opposed
to six phase space coordinates. Second, under the assumption of a phase-mixed
system, the dynamical properties of a population of stars should be uniquely a
function of their actions and independent of the conjugate angles, allowing
investigation of the relationship between \emph{orbital} properties of stars
and other intrinsic, and at least partially invariant, properties such as age
or metallicity \citep{2018ApJ...867...31B, 2018arXiv180803278T,
2018MNRAS.481.4093S, 2019arXiv190304030G, 2019arXiv190309320D,
2019MNRAS.486.1167B}. Actions also provide a convenient basis for constructing
models of the stellar distribution function
\citep[e.g.][]{1915MNRAS..76...70J, 1985ApJ...295..388V, 2017ApJ...839...61T},
or for associating stars with similar dynamical properties (e.g. to determine
membership in moving groups).

If the system being considered departs from axisymmetry in a significant
and/or non-adiabatic way, the actions computed using an axisymmetric
approximation to the true potential can exhibit cyclic dependence on the
orbital phase (or time at which the star's position and velocity are
observed), large-scale migration, or diffusion from their initial values. In
the Milky Way, stellar actions are expected to diffuse on short time scales
due to scattering with gas clouds and to evolve on longer time scales in the
case of orbits near resonances with spiral arms, bar(s), and other large scale
perturbations \citep{2014RvMP...86....1S}. For this reason, we and other
authors have used actions to study stellar scattering in the Milky Way disk
using the improved astrometry of \textit{Gaia} DR2 and various age catalogues
\citep{2018ApJ...867...31B, 2018arXiv180803278T}. Actions have also been used
to study different models of spiral structure in the Milky Way
\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in
the extended solar neighborhood in action space are discussed in
\citet{2019MNRAS.484.3291T}.

The transformation to action space is nonlinear in both the phase-space
position and the potential, and requires relatively strong assumptions about
the system's symmetry. However, the Galactic potential is not strictly
axisymmetric; with the vast improvement in the quality of phase-space
measurements due to \textit{Gaia}, this assumption appears increasingly
inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption were
close enough for many purposes, the parameters used in axisymmetric models may
be inadequately constrained by current observations.

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, their local midplane location
will not agree with our local midplane extrapolated onto their position due to
relatively small-scale variations in the density of the gas and stellar
distribution. Converting positions and velocities of more distant stars from a
heliocentric to a Galactocentric coordinate system introduces a systematic
bias in the $z$ coordinate, and hence the actions computed for these stars.
The further from the solar neighborhood the target star, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar argument applies to any remaining uncertainty
in measurements of the Galactocentric radius of the Sun and to variations in
the LSR.

\begin{figure*}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{fig/cartoon.pdf}
\end{center}
\caption{Cartoon approximation showing the effect an error in the
determination of the coordinate midplane can have on orbit integration and
action estimation. The $x$-axis shows the orbital phase and the $y$-axis the
vertical height. The top gray curve depicts an example ``true'' orbit
oscillating about the true midplane (horizontal solid gray line). Consider an
observer who erroneously assumes the midplane is located at the horizontal
dashed line. Suppose this observer measures the phase-space position of the
star at two different orbital phases (teal and orange points). If the observer
were then to integrate the star's orbit using a model potential with the
erroneous midplane, they would obtain the teal and orange curves for the
star's orbit, respectively. The actions estimated from these two  erroneous
orbits would subsequently differ, both from each other and from the actions
estimated for the true orbit (in the potential with the correct midplane).
Hence an incorrect midplane in the potential model assumed will induce
phase-dependence in the actions estimated for a given star in that potential.}
\label{fig:cartoon}
\end{figure*}

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions. We then discuss two potential
sources of these coordinate system offsets in
Sections~\ref{sec:mes_err}~and~\ref{sec:local_fire}. In
Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
parameters of the Galactocentric coordinate system. In
Section~\ref{sec:local_fire} we examine the azimuthal variations of the
midplane itself in two sets of simulations: cosmological, hydrodynamical,
zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}}
\citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a
controlled N-body simulation of a Sagittarius encounter with a galaxy
otherwise tailored to the stellar mass, scale length, and scale height of the
Milky Way \citep{2018MNRAS.481..286L}.

In Section~\ref{sec:discussion}, we discuss the implications of midplane
variations, and the resulting systematic uncertainty in the vertical action,
for action-space analyses. We also estimate the expected midplane variations
of the Milky Way based on the observed velocity variations. We summarize our
main results and conclude in Section~\ref{sec:conclusion}.

@@

\subsection{Effect of Midplane Offset} \label{ssec:cartoon}
We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
inaccurate or erroneous determination of the midplane (and hence the
zero-point of the $z$ coordinate) leads to a phase-dependence of the actions
calculated from an orbit specified by a phase-space point and an assumed
potential model. The $y$-axis corresponds to the vertical height of the orbit
and the $x$-axis shows orbital phase. The solid gray curve indicates the
``true'' orbit of the star; that is, its orbit as it oscillates around the
true midplane. The dashed gray line is the midplane location in the potential
model used by an observer to integrate the orbit of the star and hence
estimate its actions, which is offset from the true midplane of the star's
orbit. The model potential is otherwise identical to the one in which the star
is actually moving.

Now suppose this observer makes a measurement of the star's position and
velocity at the teal point or the orange point (i.e. at two different orbital
phases). Then the teal and orange curves correspond to the orbits that the
observer would compute for each point based on the potential model with the
offset midplane. In action space, this would correspond to a different value
of $J_z$ for the teal and orange points. In this way, assuming the wrong
coordinate system induces a phase-dependence in the actions estimated for the
star, which in the correct potential (in this example, the one with the
correct midplane) should be phase-independent.

\subsection{Epicyclic Approximation} \label{ssec:epi_action}

\begin{deluxetable*}{cccccccc}
\tablecaption{Description and names of the three orbits considered in this
work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of
the radial excursions of the orbit. In the epicyclic approximation, $A_z =
z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_R$} &
\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
\enddata
\end{deluxetable*}

Before turning to numerical methods, we first write down analytically the
systematic error in the actions induced from either a midplane offset or a
Galactocentric radius offset. We use the epicyclic approximation, which
assumes that the motion in the $z$ and $R$ components of the orbit are
decoupled and follow simple harmonic motion about a circular and planar
``guiding orbit'' \citep[][Section~3.2 and references
therein]{2008gady.book.....B}. We refer to the radius of this orbit (the
``guiding radius'') as $R_g$. This approximation is an excellent description
of thin disk-like orbits and a good description of thick disk-like orbits in
an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and
spiral arms). We also make the assumption of a perfectly flat circular
velocity curve with velocity $v_c$, a good approximation near the Solar circle
\citep[e.g.][]{2017MNRAS.465...76M}.

Under this approximation, we can write down the cylindrical components
of the orbits as
\beq\label{eq:orbits_epi}
\begin{split}
\phi(t) &= \Omega_c t \\
R(t) &= R_g + A_R \sin{(\kappa t + \alpha)} \\
z(t) &= A_z \sin{(\nu t + \beta)}
\text{,}
\end{split}
\eeq
where $\kappa$ and $\nu$ are the radial and vertical frequencies, $\Omega_c
\equiv v_c/R_g$ is the rotation frequency, $A_z$ and $A_R$ are the amplitudes of
the stellar motion in the vertical and radial coordinates, and $\alpha$ and
$\beta$ are arbitrary constants. Similarly, the velocities of the orbit are
given by:
\beq\label{eq:orbits_vel_epi}
\begin{split}
v_{\phi}(t) &= v_c \\
v_R(t) &= \kappa A_R \cos{(\kappa t + \alpha)} \\
v_z(t) &= \nu A_z \cos{(\nu t + \beta)}
\text{.}
\end{split}
\eeq

In this case, we have that for $J_{\phi}$
\citep[][Section~3.5.3b]{2008gady.book.....B}:
\beq\label{eq:Jphi_epi}
J_{\phi} = R_g v_c\text{,}
\eeq
and for $J_R$ and $J_z$:
\beq\label{eq:JR_Jz_epi}
\begin{split}
J_R &= \frac{E_R}{\kappa} \\
J_z &= \frac{E_z}{\nu} \text{,}
\end{split}
\eeq
where $E_R$ and $E_z$ are the energy per unit mass in the radial and vertical
coordinates, respectively. Therefore,
\beq\label{eq:JR_Jz_epi_energy}
\begin{split}
J_R &= \frac{v_R^2 + \kappa^2 (R-R_g)^2}{2\kappa} \\
J_z &= \frac{v_z^2 + \nu^2 z^2}{2\nu}\text{.}
\end{split}
\eeq
Using Equations~\eqref{eq:orbits_epi} and \eqref{eq:orbits_vel_epi}, we can
simplify this and write
\beq\label{eq:JR_Jz_epi_final}
\begin{split}
J_R &= \frac{\kappa A_R^2}{2} = \frac{v_{R,\text{max}}^2}{2\kappa} \\
J_z &= \frac{\nu A_z^2}{2} = \frac{v_{z,\text{max}}^2}{2\nu}\text{,}
\end{split}
\eeq
where the last equality in each line comes from the fact that
$v_{R,\text{max}} = \kappa A_R$ and $v_{z,\text{max}} = \nu A_z$.

>>>>

\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{bm}
\usepackage{mathtools}

@@

@@

\title{The Galactic Midplane Is Not a Plane: Implications for Dynamical
Analysis with {\em Gaia} Data and Beyond}

\correspondingauthor{Angus Beane}
\email{abeane@sas.upenn.edu}

\author{Angus Beane}
\affil{\cca}
\affil{\penn}

\author{Robyn Sanderson}
\affil{\cca}
\affil{\penn}

@@

\author{Mordecai-Mark Mac Low}
\affil{\cca}
\affil{\amnh}

@@

\author{Douglas Filho}
\affil{\columbia}

\author{David W. Hogg}
\affil{\cca}
\affil{\nyuphys}
\affil{\nyucds}
\affil{\mpia}

\author{Chervin F. P. Laporte}
\altaffiliation{Simons Fellow}
\affil{\victoria}

@@

Stellar actions, computed for stars from their six-dimensional phase space measurements and
an assumed Galactic potential, are used to label and distinguish orbits. In
principle, actions have the attractive quality of being invariant and thus
good labels of a star's orbit. 
%mm Of course,
    Such analyses are often used to interpret data from the {\em Gaia}
    mission. However, 
inaccurate measurements of the 
phase space position and uncertainties in the assumed Galactic potential
induce an error in the computed action. We show that, in addition to these
complications, a systematic bias in the phase space position induces a
phase-dependence in the actions, which we interpret as 
%mm an 
    a systematic
error. The induced error distribution is
non-Gaussian and bimodal, 
with neither of the modes peaking on
%mm  the true action value. [the error in a Gaussian peaks on zero] 
    the null value.
A variation in the vertical position midplane of $\sim15\,\pc$ for a thin disk orbit or
$\sim120\,\pc$ for a thick disk orbit is 
%mm all that is necessary 
      sufficient
to induce a $25\,\%$ systematic error in the vertical action $J_z$ computed under the assumption of an axisymmetric potential.
%mm [fragment?] orbits, respectively. 
Furthermore, we show that the local midplane varies by
$\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
simulations from the FIRE collaboration. Thus, current state-of-the-art action
calculations --- which assume the global and local midplanes are the same ---
include a systematic 
%mm offset in $z$.
   vertical offset variation. 
Variation in the local standard of rest
%mm will induce 
     induces
similar issues. The variation of the midplane must be taken into
account when performing dynamical modelling across
the large regions of the disk
accessible 
%mm with the {\em Gaia} mission. [moved first mention up to top of
%abstract, given that this is in the title.]
    to {\em Gaia}.

@@

\keywords{Galaxy:~disk -- solar~neighborhood --
open~clusters~and~associations:~general -- Galaxy:~structure --
Galaxy:~evolution}

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution as a
result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include
the remnants of a major merger \citep{2018ApJ...860L..11K,
2018Natur.563...85H, 2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space
``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly
indicating either local substructure infall \citep{2018MNRAS.481.1501B,
2018arXiv180800451L} or bar buckling \citep{2018arXiv181109205K}, and a gap
suggestive of perturbation by a dark matter subtructure in the tidal stream
GD1 \citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
indicate that the Milky Way's stellar distribution, which demonstrably departs
significantly from axisymmetry, is undergoing phase mixing and dynamical
interactions across a range of spatial and temporal scales.

@@

Once a Galactocentric coordinate system has been established and a
six-dimensional (6D) phase space measurement of a star has been made, it is
often desirable to convert this measurement into action space to concisely
summarize its projected orbit, model the stellar distribution function, or
find stars with similar dynamical properties. Actions are conserved quantities
that describe the orbit of a star under the assumption of regular, bound
orbits in a system where the equations of motion are separable in a particular
coordinate system. They are the cyclical integral of the canonical momentum
over its conjugate position,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,.}
\eeq
Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical,
and azimuthal coordinates respectively in a cylindrical coordinate system and
the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric potential, these
actions are invariant and $J_{\phi} \equiv L_z$
\citep{2008gady.book.....B,2014RvMP...86....1S}.

With the advent of 6D phase-space measurements over a relatively large
($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars'
actions has gained new popularity. One reason is dimensionality reduction---an
individual stellar orbit is concisely described by three actions, as opposed
to six phase space coordinates. Second, under the assumption of a phase-mixed
system, the dynamical properties of a population of stars should be uniquely a
function of their actions and independent of the conjugate angles, allowing
investigation of the relationship between {\em orbital} properties of stars
and other intrinsic, and at least partially invariant, properties such as age
or metallicity \citep{2018ApJ...867...31B, 2018arXiv180803278T,
2018MNRAS.481.4093S, 2019arXiv190304030G, 2019arXiv190309320D,
2019MNRAS.486.1167B}. Actions also provide a convenient basis for constructing
models of the stellar distribution function
\citep[e.g.][]{1915MNRAS..76...70J, 1985ApJ...295..388V, 2017ApJ...839...61T},
or for associating stars with similar dynamical properties (e.g. to determine
membership in moving groups).

If the system being considered departs from axisymmetry in a significant
and/or non-adiabatic way, the actions computed using an axisymmetric
approximation to the true potential can exhibit cyclic dependence on the
orbital phase (or time at which the star's position and velocity are
observed), large-scale migration, or diffusion from their initial values. In
the Milky Way, stellar actions are expected to diffuse on short time scales
due to scattering with gas clouds and to evolve on longer time scales in the
case of orbits near resonances with spiral arms, bar(s), and other large scale
perturbations \citep{2014RvMP...86....1S}. For this reason, we and other
authors have used actions to study stellar scattering in the Milky Way disk
using the improved astrometry of {\em Gaia} DR2 and various age catalogues
\citep{2018ApJ...867...31B, 2018arXiv180803278T}. Actions have also been used
to study different models of spiral structure in the Milky Way
\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in
the extended solar neighborhood in action space are discussed in
\citet{2019MNRAS.484.3291T}.

The transformation to action space is nonlinear in both the phase-space
position and the potential, and requires relatively strong assumptions about
the system's symmetry. However, the Galactic potential is not strictly
axisymmetric; with the vast improvement in the quality of phase-space
measurements due to {\em Gaia}, this assumption appears increasingly
inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption were
close enough for many purposes, the parameters used in axisymmetric models may
be inadequately constrained by current observations.

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, the local midplane location
will not agree with our local midplane extrapolated onto their position,
due to relatively small-scale variations in the density of the gas and
stellar distribution. Converting positions and velocities of more distant
stars from a heliocentric to a Galactocentric coordinate system introduces a
systematic bias in the $z$ coordinate, and hence the actions computed for
these stars. The further from the solar neighborhood the target star, the more
likely the mismatch will result in large systematic uncertainty, especially in
the vertical action $J_z$. Depending on the type of orbit being studied, this
sets a limit on the ``dynamical solar neighborhood'' over which the local
coordinate frame can be usefully extended for action-space analysis. A similar
argument applies to any remaining uncertainty in measurements of the
Galactocentric radius of the Sun and to variations in the local
standard of rest (LSR).

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions. We then discuss two potential
sources of these coordinate system offsets in
Sections~\ref{sec:mes_err}~and~\ref{sec:local_fire}. In
Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
parameters of the Galactocentric coordinate system. In
Section~\ref{sec:local_fire} we examine the azimuthal variations of the
midplane itself in two sets of simulations: cosmological, hydrodynamical,
zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}}
\citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
simulation of a Sagittarius encounter with a galaxy otherwise tailored to the stellar mass, scale length, and scale height of the Milky Way \citep{2018MNRAS.481..286L}.

In Section~\ref{sec:discussion} we discuss the implications of midplane
variations, and the resulting systematic uncertainty in the vertical action,
for action-space analyses beyond the dynamical solar neighborhood. In light of our findings, we considered the
orbits of the nine open clusters within $250\,\pc$ as reported by the {\em
Gaia} collaboration \citep{2018AA...616A..10G}. In
Table~\ref{tab:real_clusters}, we report the distance to each cluster, the
three actions, and the maximum vertical extent of their orbits
($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we estimate and
discuss the size of the dynamical solar neighborhood for each cluster based on
its mass, to show that action-space tagging is likely still a viable method if
contamination from foreground or background stars is unimportant. We summarize
and conclude in Section~\ref{sec:conclusion}.

\begin{deluxetable*}{cCCCCCCCC}
\tablecaption{Relevant dynamical parameters for the $9$ open clusters within
$250\,\pc$ of the Sun, using the relevant initial positions and velocities
from \citet{2018AA...616A..10G}. Actions are computed using the method outlined in Section~\ref{ssec:action_comp}, with typical values for the Galactocentric coordinate system parameters. These are only shown to illustrate the general types of orbits open clusters are on. The neighborhood column is explained in more
detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_R$} &
\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} \\ 
\colhead{ } & \colhead{ } &
\colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$}}
\startdata
alphaPer & 740 & 170 & 20 & -1700 & 0.005 & 11 & 352\tablenotemark{b} \\
Blanco1 & 489 & 240 & 2 & -1800 & 2 & 210 & 410\tablenotemark{c} \\
ComaBer & 153 & 90 & 2 & -1900 & 0.7 & 140 & 112\tablenotemark{d} \\
Hyades & 515 & 50 & 20 & -1800 & 0.3 & 80 & 400\tablenotemark{e} \\
IC2391 & 325 & 150 & 6 & -1800 & 0.02 & 20 & \nodata \\
IC2602 & 492 & 150 & 10 & -1700 & 0.3 & 89 & \nodata \\
NGC2451A\tablenotemark{a} & 400 & 190 & 6 & -1800 & 0.2 & 73 & \nodata \\
Pleiades & 1326 & 140 & 20 & -1700 & 0.4 & 96 & 800\tablenotemark{f} \\
Praesepe & 938 & 190 & 20 & -1800 & 0.6 & 120 & 550\tablenotemark{d}
\enddata

\tablenotetext{a}{This cluster is labeled NGC2451, which we interpret as
NGC2451A since NGC2451B lies further than $250\,\pc$ away.}
\tablenotetext{b}{\citet{2016MNRAS.457.1028S}}
\tablenotetext{c}{\citet{2007AA...471..499M}}
\tablenotetext{d}{\citet{2007AJ....134.2340K}}
\tablenotetext{e}{\citet{1998AA...331...81P}}
\tablenotetext{f}{\citet{2001AJ....121.2053A}}

\tablerefs{\citet{2018AA...616A..10G}}
\end{deluxetable*}

@@

\subsection{Effect of Midplane Offset} \label{ssec:cartoon}
We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
inaccurate or erroneous determination of the midplane (and hence the zero-point of the $z$ coordinate) leads to a phase-dependence of the
actions calculated from an orbit specified by a phase-space point and an
assumed potential model. The $y$-axis corresponds to the vertical height of
the orbit and the $x$-axis shows orbital phase. The solid gray curve indicates the
``true'' orbit of the star; that is, its orbit as it oscillates around the
true midplane. The dashed gray line is the midplane location in the
potential model used to integrate the orbit of the star and
hence estimate its actions, which is offset from the true midplane of the star's orbit. The model potential is otherwise identical to the one in which the star is actually moving.

Now suppose this observer makes a measurement of the star's position and velocity at the purple
point or the green point (i.e. at two different orbital phases). Then the
purple and green curves correspond to the orbits that the observer would
compute for each point based on the potential model with the offset
midplane. In action space, this would correspond to a different value of $J_z$
for the purple and green points. In this way, assuming the wrong coordinate
system induces a phase-dependence in the actions estimated for the star, which
in the correct potential (in this example, the one with the correct midplane)
should be phase-independent.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/cartoon.pdf}
\end{center}
\caption{Cartoon approximation showing the effect an error in the determination of the
coordinate midplane can have on orbit integration and action estimation. The
$x$-axis shows the orbital phase and the $y$-axis the vertical height. The top
gray curve depicts an example ``true'' orbit oscillating about the true
midplane (horizontal solid gray line). Consider an observer who erroneously
assumes the midplane is located at the horizontal dashed line. Suppose this
observer measures the phase-space position of the star at two different
orbital phases (purple and green points). If the observer were then to
integrate the star's orbit using a model potential with the erroneous
midplane, they would obtain the purple and green curves for the star's orbit,
respectively. The actions estimated from these two  erroneous orbits would
subsequently differ, both from each other and from the actions estimated for
the true orbit (in the potential with the correct midplane). Hence an
incorrect midplane in the potential model assumed will induce phase-dependence
in the actions estimated for a given star in that potential.}
\label{fig:cartoon}
\end{figure*}

\subsection{Epicyclic Approximation} \label{ssec:epi_action}
Before turning to numerical methods, we first write down analytically the systematic error in the actions induced from either a midplane offset or a Galactocentric radius
offset. We use the epicyclic approximation, which assumes that the motion in
the $z$ and $R$ components of the orbit are decoupled and follow simple
harmonic motion about a circular and planar ``guiding orbit'' \citep[][Section 3.2 and references therein]{2008gady.book.....B}. We refer to the radius of this orbit (the ``guiding radius'') as $R_g$. This approximation is an excellent description of
thin disk-like orbits and a good description of thick disk-like orbits in an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and spiral arms). We
also make the assumption of a perfectly flat circular velocity curve with
velocity $v_c$ \citep[decent near the Solar circle, e.g.][]{2017MNRAS.465...76M}.

Under this approximation, we can write down the cylindrical components
of the orbits as,
\beq\label{eq:orbits_epi}
\begin{split}
\phi(t) &= \Omega_c t \\
R(t) &= R_g + A_R \sin{(\nu t + \beta)} \\
z(t) &= A_z \sin{(\kappa t + \alpha)}
\text{,}
\end{split}
\eeq
where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
frequencies, $\Omega_c \equiv v_c/R_g$ the rotation frequency, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
constants. Similarly, the velocities of the orbit are given by
\beq\label{eq:orbits_vel_epi}
\begin{split}
v_{\phi}(t) &= v_c \\
v_R(t) &= A_R \nu \cos{(\nu t + \beta)} \\
v_z(t) &= A_z \kappa \cos{(\kappa t + \alpha)}
\text{.}
\end{split}
\eeq

In this case, we have that for $J_{\phi}$ \citep[][Section~3.5.3b]{2008gady.book.....B}, 
\beq\label{eq:Jphi_epi}
J_{\phi} = R_g v_c\text{,}
\eeq
and for $J_R$ and $J_z$,
\beq\label{eq:JR_Jz_epi}
\begin{split}
J_R &= \frac{E_R}{\kappa} \\
J_z &= \frac{E_z}{\nu} \text{,}
\end{split}
\eeq
where $E_R$ and $E_z$ are the energy per unit mass in the radial and vertical
coordinates, respectively. Since $a_R = -\kappa^2(R-R_g)$, and $a_z = - \nu^2
z$, we have that,
\beq\label{eq:JR_Jz_epi_energy}
\begin{split}
J_R &= \frac{\frac{1}{2}v_R^2 + \kappa^2 (R-R_g)^2}{\kappa} \\
J_z &= \frac{\frac{1}{2}v_z^2 + \nu^2 z^2}{\nu}
\end{split}
\eeq

While the formula for $J_{\phi}$ is phase independent, $J_R$ and $J_z$ are
both phase dependent via Eq. \eqref{eq:orbits_epi}. We consider two cases. First, at the point of maximum
amplitude or when $(v_R, v_z) = (0, 0)$ and second, at the point of maximum
velocity $(R, z) = (R_g, 0)$. For $J_R$ we have that,
\beq\label{eq:JR_epi_final}
    \begin{aligned}
        J_R &= \kappa A_R^2 && \text{if $v_R = 0$} \\
    J_R &= \frac{v_{R,\text{max}}^2}{2\kappa^2} && \text{if $R=R_g$} \text{.}
    \end{aligned}
\eeq
For $J_z$, we have that,
\beq\label{eq:Jz_epi_final}
    \begin{aligned}
        J_z &= \kappa A_z^2 && \text{if $v_z = 0$} \\
        J_z &= \frac{v_{z,\text{max}}^2}{2\kappa^2} && \text{if $z=0$} \text{.}
    \end{aligned}
\eeq
